Week 5  - Chapter 7 pt. 2 - Analysis

### 2024-11-07 revisitwithJochem.Rmd
checkinwithjochempush

---
title: "Plotting plots and Hotzenplotz"
author: "Verooo"
date: "2024-11-04"
output: html_document
---

@@

```
# Set-up
## Clean workspace
```{r}
rm(list = ls())
```

@@

```{r}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

@@

## Functions 
```{r}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

@@

## More functions 
```{r}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

@@

```{r}
scholars <- fload("./data/processed/scholars_20240924.rda") #make sure to lead fload first ofcourse

@@

```{r}
# save the output of your function
testsoc  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "sociology", 
                waves = list(c(2015, 2018), c(2019, 2023)),

@@

### waves - sociology department

```{r}

# Jochem messed up and did: waves, rows, columns [1, , ]
socwave1 <- testsoc$nets[1, , ]  #rows, columns, waves
socwave2 <- testsoc$nets[2, , ]  #rows, columns, waves
# let us put the diagonal to zero
diag(socwave1) <- 0
diag(socwave2) <- 0
# we want a binary tie (not a weighted tie)
socwave1[socwave1 > 1] <- 1
socwave2[socwave2 > 1] <- 1
# put the nets in an array
net_soc_array <- array(data = c(socwave1, socwave2), dim = c(dim(socwave1), 2))

str(net_soc_array)  #now we basically restructured this 
# output: num [1:50, 1:50, 1:2] 0 0 0 0 0 0 0 0 0 0

# dependent
net <- sienaDependent(net_soc_array)
```

### waves - polsci department

```{r}
# Jochem messed up and did: waves, rows, columns [1, , ]
polwave1 <- testpolsci$nets[1, , ]  #rows, columns, waves
polwave2 <- testpolsci$nets[2, , ]  #rows, columns, waves
# let us put the diagonal to zero
diag(polwave1) <- 0
diag(polwave2) <- 0
# we want a binary tie (not a weighted tie)
polwave1[polwave1 > 1] <- 1
polwave2[polwave2 > 1] <- 1
# put the nets in an array
net_pol_array <- array(data = c(polwave1, polwave2), dim = c(dim(polwave1), 2))

str(net_pol_array)  #now we basically restructured this

### My data

#### sociology dep.

not using net_soc_array for now since I am unsure about its use due to question above

@@

  # --> EGO DATAFRAME with demo and oa (sociology)
  # ?join.data.frame
  df_soc_ego <- left_join(df_ego_demo, df_ego_oa, by = join_by(au_id == id))

@@

### Gender

#### gender-soc

```{r}
df_soc_ego$Naam # list of all names in sociology department

@@

@@

#### gender - polsci

```{r}
df_pol_ego$Naam # list of all names

@@

#covariate 
# gender <- coCovar(df_ego$gender)
```

### joining soc and polsci data
```{r}
df_both_ego <- full_join(df_soc_ego, df_pol_ego, by = join_by
                    (au_id == au_id, 
                     Naam==Naam, 
                     gender == gender, 
                     discipline.22 == discipline.22,
                     discipline.24 == discipline.24,
                     counts_by_year== counts_by_year, 
                     works_count == works_count, 
                     cited_by_count == cited_by_count, 
                     Universiteit1.22 == Universiteit1.22,
                     Universiteit2.22 == Universiteit2.22,
                     Universiteit1.24 == Universiteit1.24,
                     Universiteit2.24 == Universiteit2.24,
                     Universiteit.22 == Universiteit.22, 
                     Universiteit.24 == Universiteit.24,
                     no_oa_id == no_oa_id,
                     display_name == display_name,
                     display_name_alternatives == display_name_alternatives,
                     ids== ids,
                     orcid == orcid,
                     year.22 == year.22,
                     year.24 == year.24,
                     Functie.24== Functie.24,
                     relevance_score == relevance_score,
                     affiliation_display_name == affiliation_display_name,
                     affiliation_ror == affiliation_ror,
                     affiliation_id == affiliation_id,
                     affiliation_country_code == affiliation_country_code,
                     affiliation_type == affiliation_type,
                     affiliations_other == affiliations_other,
                     affiliation_lineage == affiliation_lineage,
                     Specialisatie.22 == Specialisatie.22,
                     Specialisatie.24 == Specialisatie.24,
                     x_concepts == x_concepts,
                     works_api_url == works_api_url))
# 101 scholars (50 sociology + 51 polsci) and 43 variables (as opposed to 21 bc i could not join them all.)

?full_join

# GODDAMMIT WHY DO I ONLY FIND THIS NOW?!?!?!?!
# mydata <- sienaDataCreate(net, gender)
```

### H-index webscraping

@@

```{r}
# Add h-index and i10-index columns to df_ego
df_ego <- df_both_ego %>%
  mutate(h_index = NA_real_, i10_index = NA_real_)

@@

# Plots and such 
## wave 1 - soc
```{r}
testsoc_w1 <- igraph::graph_from_adjacency_matrix(
  testsoc$nets[1,,], # first wave of data. 
  # (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(testsoc_w1,
     vertex.label = NA,
     edge.width = 0.2,
     edge.arrow.size =0.2)

## wave 2 - soc
```{r}
testsoc_w2 <- igraph::graph_from_adjacency_matrix(
  testsoc$nets[2,,], #take the second wave 
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL
  )

plot(testsoc_w2,
     vertex.label = NA,
     edge.width = 0.2,
     edge.arrow.size =0.2)
```

## wave 1 - pol
```{r}
testpol_w1 <- igraph::graph_from_adjacency_matrix(
  testpolsci$nets[1,,], #for this example I take the first wave of data. (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
  )

plot(testpol_w1,
     vertex.label = NA,
     edge.width = 0.2,
     edge.arrow.size =0.2)

## wave 2 - pol
```{r}
testpol_w2 <- igraph::graph_from_adjacency_matrix(
  testpolsci$nets[2,,], #take the second wave 
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL
  )

plot(testpol_w2,
     vertex.label = NA,
     edge.width = 0.2,
     edge.arrow.size =0.2)
```

##  Ego characteristics 
### soc ego
```{r, include=FALSE}
#Let us find ego characteristics. 
#first fish out the data
# df <- df_ego$data

#same complicated structure as 'scholars' thus first make a dataframe from the list in which all info was saved. 
# df_ego <- do.call(rbind.data.frame, df$demographics)
#DO NOT MESS UP THE ORDER! THUS IF YOU JOIN THIS DATA WITH YOUR OWN DATA CHECK THAT ORDER REMAINED THE SAME!!

# plot(test_w2,
     # vertex.color = ifelse(df_ego$Universiteit.24 == "RU", "red", "blue"), #now, I can use actor attributes for plotting. 
     # vertex.label = NA,
     # edge.width = 0.2,
     # edge.arrow.size =0.2)
```

### pol ego
```{r, include=FALSE}
#Let us find ego characteristics. 
#first fish out the data
# df <- testpolsci$data

#same complicated structure as 'scholars' thus first make a dataframe from the list in which all info was saved. 
# df_ego <- do.call(rbind.data.frame, df$demographics)
#DO NOT MESS UP THE ORDER! THUS IF YOU JOIN THIS DATA WITH YOUR OWN DATA CHECK THAT ORDER REMAINED THE SAME!!

# plot(test_w2,
#      vertex.color = ifelse(df_ego$Universiteit.24 == "RU", "red", "blue"), #now, I can use actor attributes for plotting. 
     # vertex.label = NA,
     # edge.width = 0.2,
     # edge.arrow.size =0.2)
```

# Descriptives 
```{r}

>>>>

---
title: "Plotting plots and Hotzenplotz"
author: "Verooo"
date: "2024-11-05"
output: html_document
---

@@

Meeting with Jochem

look at publications website (sunbelt workshop)

# Getting started

```{r clean}
rm(list = ls())
```

@@

```{r package}
require(tidyverse)
require(RSiena)
require(RsienaTwoStep)

@@

## Functions

```{r functions 1}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

@@

## More functions

```{r functions 2}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology"+"political science", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

@@

```{r data}
scholars <- fload("./data/processed/scholars_20240924.rda") #make sure to lead fload first ofcourse
```

```{r saving output}
# save the output of your function

test  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = c("sociology", "political science"),
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("all")) # all, first, last # which ties in there

# USE not all but first or last so the ties are directed from "perspective" of first author of the publications 
# usually first most prestige

@@

# Gender

## saving data for each deapartment

```{r}
#inspecting scholars_20240924.rda object
#df_ego <- do.call(rbind.data.frame, scholars$demographics)
#unique(df_ego$discipline.22)

# save the output of your function
testsoc  <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "sociology", 
                waves = list(c(2015, 2018), c(2019, 2023)),

@@

#### soc department

@@

  # --> EGO DATAFRAME with demo and oa (sociology)
  ?join.data.frame
  df_soc_ego <- left_join(df_ego_demo, df_ego_oa, by = join_by(au_id == id))

@@

### gender-soc

```{r gender s}
df_soc_ego$Naam # list of all names in sociology department

@@

# fsave(df_soc_ego)

@@

### gender - polsci

```{r gender p}
df_pol_ego$Naam # list of all names

@@

#covariate 
# gender <- coCovar(df_ego$gender)

# fsave(df_pol_ego)
```

### save gender data sets

```{r gender data}
soc_gender <- fload("./data/processed/df_soc_ego_20241105.rda")
pol_gender <- fload("./data/processed/df_pol_ego_20241105.rda")
```

### Combine data (gender)

```{r combine gender set}
df <- test$data
df_ego <- do.call(rbind.data.frame, df$demographics)

soc_gender <- soc_gender[,c("Naam", "gender")]
pol_gender <- pol_gender[,c("Naam", "gender")]

gender <- rbind(soc_gender, pol_gender) # combining data sets

df_ego <- df_ego %>% 
            left_join(gender) #dataset 
```

# Quick RSiena

```{r RSiena}
#step1 data
wave1 <- test$nets[1,,]
wave2 <- test$nets[2,,]

#put diagonals to zero 
diag(wave1) <- 0
diag(wave2) <- 0
# we want a binary tie (not a weighted tie)
wave1[wave1 > 1] <- 1
wave2[wave2 > 1] <- 1
# put the nets in an array
net_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2))

gender <- df_ego$gender

#dependent
nets <- sienaDependent(net_array)
#covariate
genderrs <- coCovar(gender)
#define data 
mydata <- sienaDataCreate(nets, genderrs)

#define effects structure 
myeff <- getEffects(mydata)

#skipped description

myeff <- includeEffects(myeff, isolateNet, inPop, outAct)

myeff <- includeEffects(myeff, egoX, altX, sameX, interaction1 = "genderrs")

#look up which effects to iclude

myAlgorithm <- sienaAlgorithmCreate(modelType = c(nets=6), projname = "proProject")
# undirected: model 6 "additive model", #check manual for this

# estimate
model1 <- siena07(myAlgorithm, data = mydata, effects = myeff, returnDeps = TRUE)
#latter needed for GOF

model1

# output:  convergence ration NA probably bc we chose all ties to be included (type = "all") earlier
# we can go back and change that or change it here in the myAlgorithm function

# Make it a directed network 
# focus: writing, h index, 
# model interpretation

# H-index webscraping

@@

```{r h-index, include=FALSE, eval= FALSE}
# Add h-index and i10-index columns to df_ego
df_ego <- df_both_ego %>%
  mutate(h_index = NA_real_, i10_index = NA_real_)

@@

# Plotting

```{r}
test_w1 <- igraph::graph_from_adjacency_matrix(
  test$nets[1,,], # first wave of data. 
  # (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL)

plot(test_w1,
     vertex.label = NA,
     edge.width = 0.2,
     edge.arrow.size =0.2)
```

```{r}
names(df_ego) #see labels of variables

plot(test_w1,
vertex.color = ifelse(df_ego$gender == 1, "red", "blue"), #now, I can use actor attributes for plotting.
vertex.label = NA,
edge.width = 0.2,
edge.arrow.size =0.2)

# if you dont want plot to turn, save the layout (look at SNASS)
```

### 2024-11-07 week5-2.Rmd
checkinwithjochempush

```
THIS IS ALL THE EFFECTS YOU COULD LOOK AT IN YOUR NETWORK 
e.g. outdegree (density), called density, three things: evaluation, endowment, and creation
  you can only include two of the three (tie creation, breaking and maintaining)
  Assumption meachnisms to make and breka a tie are the same so we only use the EVALUATION part

  All these effects are describedin the manual of RSiena Chapter 12 (9?)

@@

# STUDY CHAPTER 12 of RSiena Manual! 
# COLLECTIVELY SEND PAPER TO JOCHEM BEFORE WEEKEND!
# PrACTICE WITH RSIENA 
# LOOK AT MANUAL

>>>>

```
THIS IS ALL THE EFFECTS YOU COULD LOOK AT IN YOUR NETWORK 
e.g. outdegree (density), called density, three things: evaluation, endowment, and creation
  you can only include two of the three (tie creation, breaking and maintaining)
  Assumption mechanisms to make and breaka a tie are the same so we only use the EVALUATION part

  All these effects are described in the manual of RSiena Chapter 12 (9?)

@@

# STUDY CHAPTER 12 of RSiena Manual! 
# PrACTICE WITH RSIENA
